---
title: "Life expectancy in Iran over time"
author: "Jaeda Yurkiw"
date: 2026-09-23
---

In this post I use R to explore how life in Iran changed between 1952 and 2007.

Data: [gapminder R package](https://cran.r-project.org/package=gapminder)
by Jennifer Bryan (CC0), an excerpt of data from the
[Gapminder Foundation](https://www.gapminder.org/data/) (CC BY 4.0).

## Setup

```{r}
#| label: setup
#| message: false
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
```

## Removing missing life expentancy values

```{r}
countries <- gapminder |>
    filter(!is.na(lifeExp))
head(countries)
```


## Life expectancy for Iran

```{r}
iran_life_exp <- countries |>
    filter(country == "Iran") |>
    select(year, lifeExp)
head(iran_life_exp)
```


## Life expectancy in descending order 

```{r}
iran_life_exp_desc <- iran_life_exp |>
    arrange(desc(lifeExp))
head(iran_life_exp_desc)
```

## Visualizing life expectancy for Iran

```{r}
#| label: fig-iran-life-exp
#| fig-cap: "Life expectancy in Iran, 1952 to 2007."
ggplot(iran_life_exp, aes(x = year, y = lifeExp)) +
  geom_line() +
  geom_point() +
  labs(x = "Year", y = "Life expectancy (years)")
```

## What I did and what I found 

I first removed rows with missing life expectancy values, then filtered the data to include only Iran. Next, I sorted the years by life expectancy in descending order to see the highest and lowest, and finally plotted life expectancy over time as a line chart.

Life expectancy in Iran increased steadily from 1952 to 2007. In 1952, life expectancy was around 50 years, and by 2007 it had risen to around 72 years, an increase of about 22 years (@fig-iran-life-exp).